Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2013-12-13 ms.tex
more text, more references

\newcommand{\IRAS}{{\it IRAS}}
\newcommand{\HERSCHEL}{{\it Herschel}}
\newcommand{\SPITZER}{{\it Spitzer}}

@@

\shorttitle{Planck Dust Model}

@@

\title{Two-component thermal dust emission model: application to the Planck HFI maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

\begin{abstract}
We have fit the \cite{FDS99} two-component dust emission model over
the full sky, based on Galactic dust SEDs composed of  Planck HFI and \cite{SFD} 
\IRAS~100$\mu$m. As a result, we have obtained maps of dust optical depth $\tau_{545}$ 
and temperatures $T_1$ and $T_2$,  with angular resolution  $\sim$6$'$. These maps 
serve as an interesting comparison to alternative models, for example 
one-component models with varying $T$ and $\beta$. We also expect the derived 
optical depth to be useful as a cross-check for other extinction estimates, 
such as the SFD98 map, which assumed $\beta = 2$ single-component dust 
emission, as well as forthcoming dust maps based on observations of stars in 
the optical rather than infrared emission. %mention the enhanced temperature resolution relative
% to SFD
\end{abstract}

\section{Introduction}

@@

We have made use of the Planck 2013 data release, specifically the 217 GHz, 
353 GHz, 545 GHz and 857 GHz channels Planck Collaboration (2013a). We use the 
versions of these maps which have been corrected for zodiacal light by the 
Planck team, using the zodiacal light model described in Planck Collaboration 
(2013b). For each pixel on the sky, Planck thus provides us with four data 
points of the dust emission SED. The angular resolution of these maps is 
$5.5$. TODO: actually smooth all of the maps to a common PSF of XX arcmin.

To add information about the dust emission on the Wien side of its spectrum,
we include IRAS 100$\mu$m. We use the SFD reprocessing of IRAS 100$\mu$m, which
we will refer to as \verb|i100| (Schegel et al. 1998). \verb|i100|

@@

@@

\subsection{Zodiacal Light Removal}
Cite the appropriate reference for the Planck ZL model.

\subsection{Smoothing to common resolution}

\subsection{CMB Subtraction}
The CMB is effectively imperceptible at 857 GHz, but can be noticed at a 
low level at 545 GHz, and is extremely bright at 217 GHz and 353 GHz relative 
to the diffuse Galactic emission we wish to characterize.

@@

\subsection{Correlation against HI}
The raw Planck maps we have downloaded show offsets relative to HI, in the
sense that the FIR emission does not go to zero when HI emission goes to
zero. Since HI correlates linearly with far-IR emission in the diffuse ISM, 
when the optical depth is low and molecular effects are negligible, we modeled
far IR emission as a constant times HI emission, plus some offset. We then
subtracted the best-fit offset such that

@@

\section{Model}
Our model is that of FDS99, which consists of two populations of dust 
grain emitting as modified blackbodies, but with different temperatures
and different emissivity power law indices $\beta$.

@@

\begin{figure*}[ht]
\begin{center}
\epsfig{file=t2_0128.eps, width=7.0in}
\caption{Our best-fit $T_2$, binned to 27.5$'$ resolution}
\end{center}
\end{figure*}

>>>>

\newcommand{\IRAS}{{\it IRAS}}
\newcommand{\HERSCHEL}{{\it Herschel}}
\newcommand{\SPITZER}{{\it Spitzer}}
\newcommand{\PLANCK}{{\it Planck}}

@@

\shorttitle{\PLANCK~Dust Model}

@@

\title{Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

\begin{abstract}
We present full-sky maps of dust optical depth and temperature derived from 
the \cite{FDS99} two-component dust emission model fitted to \PLANCK~217, 353, 
545, 857 GHz and \cite{SFD} \IRAS~100$\mu$m data. Both the optical 
depth and temperature maps have angular resolution of $6.1'$. Such maps serve 
as an important alternative to and comparison for other dust 
emission models, especially the single modified blackbody model in which 
$\beta$ and $T$ both vary spatially. We also expect the derived optical 
depth to provide valuable cross-checks for other dust-column related data 
products, specifically the \cite{SFD} map, which assumed 
$\beta = 2$ single-component dust emission and included a temperature map of 
inferior $\sim$$1.3^{\circ}$ resolution. We briefly compare our 
best-fit two-component spectral energy distributions to those of the 
\cite{planckdust} single-component parametrization.
%mention the enhanced temperature resolution relative to SFD
%Our optical depth map will also provide a useful reference for dust maps 
%based on stellar colors rather than infrared emission.
\end{abstract}

\section{Introduction}
%why dust important in astronomy/astrophysics
The presence of Galactic interstellar dust affects
astronomical observations over a wide range of wavelengths. At mid-infrared
and far-infrared wavelengths, Galactic dust emission is a major contributor
to the observed sky intensity. In the optical and ultraviolet, extinction
by foreground dust grains attenuates the signal from extragalactic sources
over the entire sky. Understanding the properties of dust in the interstellar 
medium (ISM) is valuable in its own right, as the ISM is the site of star 
formation, thereby playing a crucial role in galaxy evolution. Equally if not 
more important to the practice of astronomy, however, is accounting for dust 
as a foreground which reddens optical/UV observations and superimposes 
Galactic emission on low-frequency observations of the 
cosmic microwave background (CMB). %specify ``low frequency more precisely eventually

%basic background about interstellar dust emission

%maybe say instead ``over the past 30/40 years...''
%Barnard first recognized evidence for dust in astronomical observations, and
%ever since it has been mapped in increasing detail. 
Over the past decades, satellite observations have dramatically enhanced our
knowledge about infrared emission from the ISM. High-latitude complexes of 
resolved infrared sources were first identified by \cite{low84} using 
\IRAS~60$\mu$m and 100$\mu$m observations \citep{wheelock94}. This so-called 
``infrared cirrus,'' attributable to thermal emission from large dust grains in
the interstellar medium, has since been probed in greater detail by a variety 
of instruments and detected over virtually the entire sky. Beginning in 1989, 
DIRBE  mapped the full sky at ten infrared wavelengths from 1.25$\mu$m to 
240$\mu$m with a reliable zero point, but inferior 0.7$^{\circ}$ angular 
resolution \citep{boggess92}.mention FIRAS specifically, since it is important 
to FDS99.

FDS99 used FIRAS to make a globally best-fit model of dust emission 
appropriate over a very broad range of frequencies. They showed that a 
model consisting of a single modified blackbody with power law index $\beta$ 
could not fit the IRAS/FIRAS/DIRBE spectrum at both the Wien and 
Rayleigh-Jeans extremes. To fit both ends of the spectrum, FDS99 proposed
a spectrum consisting of two modified black bodies, each with a different
temperature and power law index. Physically, these two components
represent distinct dust grain species within the ISM, specifically 
silicates and carbonaceous grains.

%how what we're doing fits in
%     where FDS99 fit in
%     Planck data qualities that improve on data used in FDS99
%     benefits of using Planck + FDS99 model
%utility/applications of what we're doing

%list of figures : 
% posterior surface for some pixel, with samples from markov chain overplotted
%    --> also overplot output best-fit (T,I545) and unctertainty estimate
% plot for some set of pixels of SED overlaid with best-fit two-component model
%    --> done in Python instead of IDL
%    --> overlay 
% plot of resulting T2 and tau
% plots of full-sky residuals relative to our model ?
%  ---> and for lower-frequency bands to emphasize that FDS99 is better for
%       extrapolating towards mm wavelengths

@@

All \PLANCK~data products utilized throughout this work are drawn from the\PLANCK~2013 data release \citep{planck2013}. Specifically, we have made use of the
the 217 GHz, 353 GHz, 545 GHz and 857 GHz intensity maps. We use the 
versions of these maps which have been corrected for zodiacal light 
\citep[\texttt{R1.10\_nominal\_ZodiCorrected},][]{planckzodi}.

%For each pixel on the sky, Planck thus provides us with four data points of 
%the dust emission SED. The angular resolution of these maps is $5.5$.

%TODO: actually smooth all of the maps to a common PSF of 6.1 arcmin.

To incorporate information about the dust emission on the Wien side of its 
spectrum, we make use of 100$\mu$m data. In particular, we use the SFD 
reprocessing of IRAS 100$\mu$m, which we will refer to as \verb|i100| (Schegel et al. 1998). \verb|i100| has angular resolution of $6.1'$. %list other
%important characteristics of i100 e.g. that it has been zodi-corrected
% and relies on dirbe on scales larger than ???

@@

\subsection{Smoothing}

@@

%\subsection{Zodiacal Light}
%As mentioned above, we rely upon the \cite{planckzodi} zodiacal light 
%correction.

\subsection{CMB Removal}
The CMB is effectively imperceptible at 857 GHz, but can be noticed at a 
low level at 545 GHz, and is prominent at 217 GHz and 353 GHz relative 
to the diffuse Galactic emission we wish to characterize, especially at high
latitudes. To remove the CMB anisotropies, we have subtracted the SMICA model
from each of our Planck maps \citep{smica}, after appropriate unit conversions.

@@

\subsection{Correlation against HI}
The raw \PLANCK~maps we have downloaded show offsets relative to HI, in the
sense that the FIR emission does not go to zero when HI emission goes to
zero. Since HI correlates linearly with far-IR emission in the diffuse ISM, 
when the optical depth is low and molecular effects are negligible, we modeled
far IR emission as a constant times HI emission, plus some offset. We then
subtracted the best-fit offset such that

@@

\section{Dust Emission Model}
Our model is that of FDS99, which consists of two populations of dust 
grains emitting as modified blackbodies, but with different temperatures
and different emissivity power law indices $\beta$. The shape of this spectrum 
is given by:

\begin{equation}
I_{\nu} \propto f_{1}q_{1}(\nu/\nu_{0})^{\beta_1}B_{\nu}(T_1) + (1-f_{1})q_{2}(\nu/\nu_0)^{\beta_2}B_{\nu}(T_2)
\end{equation}

With $\nu_0$ = 3000 GHz, $f_{1}$ = 0.0363, $\beta_1$ = 1.67, $\beta_2$ = 2.70, 
and $q_1/q_2$ = 13.0. For the parameters thus far specified, the two dust temperatures are further related by:

\begin{equation}
T_1 = 0.352T_2^{1.18}
\end{equation}

At each sky location, we have five independent intensity 
measurements (four ) FDS99 found typical values for T2 and

@@

\section{Pixelization}
For the purpose of fitting, we break the sky into independent pixels of 
angular diameter $\sim$, defined by the HEALPix pixelization with nside=2048. 
This is convenient because the Planck maps are provided in this format and 
this pixelization adequately samples the $6.1'$ resolution maps under 
consideration. We reiterate that each pixel is fit independently of 
neighboring pixels.

\section{Markov Chains}
As explained in $\S$ %need to label ``model'' section so i can reference it

\section{Results}

\subsection{Calculating $\tau$}

\subsection{Data Release}
We are releasing nside=2048 HEALPix maps in Galactic coordinates summarizing 
the results of of our two-component dust fits.

\begin{figure*}
\begin{center}
\epsfig{file=results.eps, width=7.0in}
\caption{Our best-fit $T_2$, binned to 27.5$'$ resolution}
\end{center}
\end{figure*}
